Refinements for open automata

### 2023-04-24 2021-Refinement/paper/refinementarticleV2.tex
examples

\begin{example}[\texttt{prod-cons}]\label{ex:fifo-system} As a running example, we consider a classical producer-consumer pair  interacting through FIFO buffer,  referred to as \texttt{prod-cons}.  Fig. \ref{Fig:Architect} reflects the overall structure of the system involving a producer process, a consumer process and an orchestrator that coordinates  their activities.

@@

\begin{example} 
 Fig. \ref{Fig:procandcons} shows a producer automaton  and a consumer  automaton that can be used  to fill the holes $\texttt{P}$ and $\texttt{Q}$ of \texttt{prod-cons} defined in Example~\ref{ex:fifo-system}.
 \begin{figure}[!ht]
 \centering
   \includegraphics[width=.35\textwidth]{Figures/P-proc.pdf}\hfill 
   \includegraphics[width=.45\textwidth]{Figures/Q-proc.pdf}
   \caption{(Left) A \texttt{producer}.  It produces one item at a time and pushes it.  (Right)  A \texttt{consumer}. It pops an item, does some work and pushes the result. \label{Fig:procandcons}}
\end{figure}

@@

\begin{figure}[!ht]
 \centering
   \includegraphics[width=.99\textwidth]{Figures/compos-FIFO-Pro.pdf}
   \caption{Open Automaton of the \texttt{prod-cons}[\texttt{P}/\texttt{producer}]. 
   \label{Fig:ComposeOA}} 
\end{figure}
\end{example}

@@

An composable automaton cannot transform external non-determinism of the sub-automaton into internal non-determinism.

@@

\begin{example} 
To illustrate the refinement simulation of open automata, consider a variation on the \texttt{prod-cons}, referred to as \texttt{Simpleprod-cons}.  We suppose that  the two processes  \texttt{P} and  \texttt{Q} communicate through a one-place  buffer. The  corresponding behaviour  is given by  an open automaton  depicted in Fig. \ref{Fig:SpecOA}. 
\begin{figure}[!t]
 \centering
   \includegraphics[width=.90\textwidth]{Figures/FIFOpen2.pdf}
   \caption{Open Automation of the \texttt{prod-cons} system using one-place  buffer.  \label{Fig:SpecOA}}
\end{figure}
We can easily check that this automaton simulates the one on Fig.~\ref{Fig:RefineOA}.
The simulation relation is $R=\{(r_0,s_0) \mapsto \texttt{l}=\texttt{f}, (r_0,s_1) \mapsto \texttt{f}=\texttt{l}+1\% \texttt{N} \}$ is a simulation. \qed
\end{example}

@@

\begin{theorem}[Congruence]
Let $A_1$, $A_2$ and $A_3$ be three open automata with $\wrel{A_2}{A_3}{H}$. 
  Let $J_3$ be the set of holes of $A_3$ and suppose that \(k \in H\). Suppose additionally that the composition $A_1\subst{A_2}{k}$ is non-blocking.
We have: \[\wrel{A_1\subst{A_2}{k}}{A_1\subst{A_3}{k}}{J_1 \uplus H \setminus \mbrc{k}}\]
\end{theorem}

\begin{example}

\texttt{prod-cons[P/producer] < simple-prod-cons[P/producer]}

The composition of the refined \texttt{database} and the \texttt{counter} is again an open automata, depicted in Fig.~\ref{Fig:ComposeRefine}. As stated by the theorem,  it can easily be verified that this automaton is a refinement of the automaton of the figure Fig.~\ref{Def:CompOA}, i.e. the automaton resulting from the composition of the abstract \texttt{database}.
\end{example}

>>>>

\begin{example}[\texttt{prod-cons}]\label{Example:fifo-system} As a running example, we consider a classical producer-consumer pair  interacting through FIFO buffer,  referred to as \texttt{prod-cons}.  Fig. \ref{Fig:Architect} reflects the overall structure of the system involving a producer process, a consumer process and an orchestrator that coordinates  their activities.

@@

\begin{example} \label{Example:prodandcons} Fig. \ref{Fig:procandcons} shows a producer automaton  and a consumer  automaton that can be used  to fill the holes $\texttt{P}$ and $\texttt{Q}$ of \texttt{prod-cons} defined in Example~\ref{ex:fifo-system}.
 \begin{figure}[!ht]
 \centering
   \includegraphics[width=.35\textwidth]{Figures/P-proc.pdf}\hfill 
   \includegraphics[width=.45\textwidth]{Figures/Q-proc.pdf}
   \caption{(Left) A \texttt{producer}.  It produces one item at a time and pushes it.  (Right)  A \texttt{consumer}. It pops an item, does some work and pushes the result. \label{Fig:procandcons}}
\end{figure}

@@

\begin{figure}[!ht]
 \centering
   \includegraphics[width=.99\textwidth]{Figures/compos-FIFO-Pro.pdf}
   \caption{Open Automaton for \texttt{prod-cons}[\texttt{P}/\texttt{producer}]. 
   \label{Fig:ComposeOA}} 
\end{figure}
\end{example}

@@

A composable automaton cannot transform external non-determinism of the sub-automaton into internal non-determinism.

@@

\begin{example} \label{example:fifo-system-one}
To illustrate the refinement simulation of open automata, we consider a variation on the \texttt{prod-cons} example. Namely,  we suppose that  the two processes  \texttt{P} and  \texttt{Q} communicate through a one-place  buffer. The  corresponding behaviour of this simpler version of the system is specified by the    open automaton  depicted in Fig. \ref{Fig:SpecOA}, referred to as \texttt{Simprod-cons}. 
\begin{figure}[!t]
 \centering
   \includegraphics[width=.90\textwidth]{Figures/FIFOpen2.pdf}
   \caption{  The \texttt{Simprod-cons} Open Automaton: \texttt{prod-cons} system using one-place  buffer.  \label{Fig:SpecOA}}
\end{figure}
We can easily check that this automaton simulates the one on Fig.~\ref{Fig:RefineOA}.
The simulation relation is $R=\{(r_0,s_0) \mapsto \texttt{l}=\texttt{f}, (r_0,s_1) \mapsto \texttt{f}=\texttt{l}+1\% \texttt{N} \}$ is a simulation. \qed
\end{example}

@@

\begin{example}  Consider again the \texttt{prod-cons}  \texttt{Simprod-cons}  automata given in examples above.  Since we have \texttt{prod-cons} $\leq_{\{\texttt{P,Q}\}}$ \texttt{simprod-cons},  hence according to Theorem \ref{thm:ContextRefinement}, we get 
\texttt{prod-cons[P/producer]} $\leq_{\{\texttt{Q}\}}$  \texttt{Simprod-cons[P/producer]}. This can be verified by comparing the behaviour of the automaton of the figure Fig.~\ref{Fig:ComposeOA} and the automaton resulting  from the composition of the \texttt{Simprod} and  \texttt{producer} automata.\qed
\end{example}

\begin{theorem}[Congruence]\label{thm:CongurenceRefinement}
Let $A_1$, $A_2$ and $A_3$ be three open automata with $\wrel{A_2}{A_3}{H}$. 
  Let $J_3$ be the set of holes of $A_3$ and suppose that \(k \in H\). Suppose additionally that the composition $A_1\subst{A_2}{k}$ is non-blocking.
We have: \[\wrel{A_1\subst{A_2}{k}}{A_1\subst{A_3}{k}}{J_1 \uplus H \setminus \mbrc{k}}\]
\end{theorem}

As an example of the use of this theorem, consider a refined version of the producer process of Example ~\ref{Example:prodandcons} that we  refer to as $\texttt{producer}^{*}$. According to  Theorem \ref{thm:CongurenceRefinement}, we have 
\texttt{prod-cons[P/producer]} $\leq_{\{\texttt{Q}\}}$  \texttt{prod-cons[P/$\texttt{producer}^{*}$]}; as well as \texttt{Simprod-cons[P/producer]} $\leq_{\{\texttt{Q}\}}$  \texttt{Simprod-cons[P/$\texttt{producer}^{*}$]}.
